An Embedding Framework for Consistent Polyhedral Surrogates

### 2021-06-23 poly-embed-journal/poly-embed-journal.tex
sec 6; 2 proofs to go

\raf{Theme: expanding our results from polyhedral losses to polyhedral Bayes risks. We can start by observing that the latter is a weaker condition.}

\raf{Moved}
(We say a concave function is polyhedral if its negation is a polyhedral convex function.)
Note that the Bayes risk, a function from distributions over $\Y$ to the reals, may be polyhedral even if the loss itself is not.
For example, this can be seen by a loss that is polyhedral in the convex hull of $\varphi(\R)$, but smooth outside that region.

\begin{lemma}
	\label{lem:poly-loss-poly-risk}
	If $L$ is polyhedral, $\risk{L}$ is polyhedral.
\end{lemma}
\begin{proof}
	Let $L:\reals^d\to\reals_+^\Y$ be a polyhedral loss, and $\Gamma = \prop{L}$.
	By Lemma~\ref{lem:polyhedral-range-gamma}, $\U = \Gamma(\simplex)$ is finite. 
	For each $U\in \U$, select $u_U \in U$, and let $\Sc = \{u_U : U \in\U\}$.
	$\Sc$ is representative for $L$, so Lemma~\ref{lem:loss-restrict} gives us $\risk{L} = \risk{L|_{\Sc}}$, which is polyhedral as $\Sc$ is finite.
\end{proof}

\jessie{Lemma X}
\begin{lemma}\label{lem:X}
  Let $L: \R \to \reals^\Y_+$ be a minimizable loss with a polyhedral Bayes risk $\risk L$.
  Then $L$ has a finite minimum representative set.
  Furthermore, letting $\Gamma = \prop{L}$, there exist finite sets
  $\V \subseteq \reals^\Y_+$ and
  $\Theta = \{\theta_v \subseteq \simplex \mid v\in\V\}$,
  both uniquely determined by $\risk{L}$ alone,
  such that
  \begin{enumerate}
  % \item There exists $\V = \V(\risk{L}) \subseteq \reals^\Y_+$, uniquely determined by $\risk{L}$ alone, such that $\V = L(\R^*)$.\label{item:X-V}
  \item A set $\R'\subseteq\R$ is representative if and only if $\V \subseteq L(\R')$.\label{item:X-rep-V}
  \item A set $\R'\subseteq\R$ is minimum representative if and only if $L(\R') = \V$.\label{item:X-min-V}
  \item A set $\R'\subseteq\R$ is representative if and only if $\Theta \subseteq \{\Gamma_r \mid r \in \R'\}$.\label{item:X-rep-Theta}
  \item A set $\R'\subseteq\R$ is minimum representative if and only if $\{\Gamma_r \mid r \in \R'\} = \Theta$.\label{item:X-min-Theta}
  \item Every representative set for $L$ contains a minimum representative set for $L$.\label{item:X-rep-contain-min}
  \item The set of full-dimensional level sets of $\Gamma$ is exactly $\Theta$.\label{item:X-full-dim}
  \item For any $r \in \R$, there exists $\theta \in \Theta$ such that $\Gamma_r \subseteq \theta$.\label{item:X-redundant}
  \item $L$ tightly embeds $\ell:\R'\to\reals^\Y_+$ if and only if $\ell$ is injective and $\ell(\R') = \V$.
  \end{enumerate}
\end{lemma}

@@

Previous work from~\citet[Proposition 4]{duchi2018multiclass} realized the significance of matching Bayes risks for calibration with respect to the 0-1 loss, but their result relies the Bayes risk of the surrogate being strictly concave.  As polyhedral functions are never strictly concave, Proposition~\ref{prop:embed-bayes-risk} extends their result to the embedding setting.

@@

Motivated by this observation, let us define the ``trim'' of a loss to be this unique set of loss vectors induced by any minimum representative set.

\begin{definition}[Trim]\label{def:trim-loss}
  Given a loss $L:\R \to \reals_+^\Y$ with a finite representative set, we define $\trimcover(L) = \{L(r) \mid r \in \R'\}$ given any minimum representative set $\R'$ for $L$.
\end{definition}

Using this notion of trimming a loss, we can again recast our embedding condition: a loss embeds another if and only if they have the same $\trim$.

@@

\begin{corollary}\label{prop:trim-loss-condition}
  Let $L : \R \to \reals^\Y_+$ have a finite representative set.
  Then $\trimcover(L) = \{L(u) \mid u \in \R \textrm{ s.t. } \neg\exists u'\in\R,u'\neq u,\, \Gamma_u \subsetneq \Gamma_{u'}\}$. 
\end{corollary}
\begin{proof}
  \raf{Didn't get to this proof}

@@

As we define one loss embedding another, one can imagine an analogous definition where one property embeds another (Definition~\ref{def:prop-embed}), and we can define $\trimcover(\Gamma)$ as the set of level sets generated by a minimum representative set (Definition~\ref{def:trim-prop}).
See \S~\ref{app:embed-props} for precise definitions and discussion.
Even when using the language of properties, similar results to those above hold; for intuition, the level sets of a property correspond to the subgradients of the Bayes risk of the loss eliciting it.
Since embedding also means matching Bayes risks, the matching of $\trimred$ follows since the subgradients also trivially match.

We now state a useful result for proving the existence of an embedding loss, which shows remarkable structure of embeddable properties, and the properties that embed them.
First, we conclude that any embeddable property must be elicitable.
We also conclude that if $\Gamma$ embeds $\gamma$, the level sets of $\Gamma$ must all be redundant relative to $\gamma$.
In other words, $\Gamma$ is exactly the property $\gamma$, just with other reports filling in the gaps between the embedded reports of $\gamma$.
(When working with convex losses, these extra reports are typically the convex hull of the embedded reports.)
In this sense, we can regard embedding as only a slight departure from direct elicitation: if a loss $L$ elicits $\Gamma$ which embeds $\gamma$, we can think of $L$ as essentially eliciting $\gamma$ itself.
Finally, we have an important converse: if $\Gamma$ has finitely many full-dimensional level sets, or if $\trimred(\Gamma)$ is finite, then $\Gamma$ must embed some finite elicitable property with those same level sets.

\raf{Moved}

\raf{Would suggest $r$ not $u$ here.}
Define $\trimred(\Gamma) = \{\Gamma_u \mid u \in \R \textrm{ s.t. } \neg\exists u'\in\R,u'\neq u,\, \Gamma_u \subsetneq \Gamma_{u'}\}$.

\begin{restatable}{proposition}{embedtrim}\label{prop:embed-trim}
  Let $\Gamma:\simplex\toto\reals^d$ be an elicitable property.
  The following are equivalent:
  \begin{enumerate}\setlength{\itemsep}{0pt}
  \item $\Gamma$ embeds a elicitable finite property $\gamma:\simplex \toto \R$.
  \item $\trimred(\Gamma)$ is a finite set.%, and $\cup\,\trimred(\Gamma) = \simplex$.
  % \\ $\hat\Gamma := \trimred(\Gamma)$ is a finite property, and $\cup\,\hat\Gamma) = \simplex$.
  \item There is a finite minimum representative set $\U$ for $\Gamma$.
  \item There is a finite set of full-dimensional level sets $\Theta$ of $\Gamma$, and $\cup\,\Theta = \simplex$.
  \end{enumerate}
  % \raft{I see why you wanted this addition here, but (a) it's clear from the def of embedding, and (b) we don't have varphi defined.  So let's leave it out.}
  Moreover, when any of the above hold, $\trimred(\gamma) = \trimred(\Gamma) = \{\Gamma_u \mid u\in\U\} = \Theta$.
\end{restatable}
\begin{proof}
  \raf{Now using Lemma~\ref{lem:X}}
\end{proof}

>>>>

While we have focused on polyhedral losses thus far, it turns out that many of our results about embeddings extend to losses with polyhedral Bayes risks.
(We say a concave function is polyhedral if its negation is a polyhedral convex function.)
To see that this is a weaker condition, recall that Theorem~\ref{thm:poly-embeds-discrete} constructs a finite representative set $\Sc$ for any polyhedral loss $L$, and thus $\risk{L} = \risk{L|_\Sc}$ by Lemma~\ref{lem:loss-restrict}, which is polyhedral.
The condition is strictly weaker, in that the Bayes risk may be polyhedral even if the loss itself is not.
For example, a modified hinge loss $L(r)_y = \max(r^2-1,1-ry)$, which matches hinge loss on the interval $[-1,1]$ but is smooth outside the interval $[-2,2]$, still embeds twice 0-1 loss.

% \begin{lemma}
% 	\label{lem:poly-loss-poly-risk}
% 	If $L$ is polyhedral, $\risk{L}$ is polyhedral.
% \end{lemma}
% \begin{proof}
% 	Let $L:\reals^d\to\reals_+^\Y$ be a polyhedral loss, and $\Gamma = \prop{L}$.
% 	By Lemma~\ref{lem:polyhedral-range-gamma}, $\U = \Gamma(\simplex)$ is finite. 
% 	For each $U\in \U$, select $u_U \in U$, and let $\Sc = \{u_U : U \in\U\}$.
% 	$\Sc$ is representative for $L$, so Lemma~\ref{lem:loss-restrict} gives us $\risk{L} = \risk{L|_{\Sc}}$, which is polyhedral as $\Sc$ is finite.
% \end{proof}

We now present our main structural result, which will lay the foundation for the rest of this section.
\raf{TODO: Brief description of the result}

\begin{lemma}[\jessie{Lemma X}]\label{lem:X}
  Let $L: \R \to \reals^\Y_+$ be a minimizable loss with a polyhedral Bayes risk $\risk L$.
  Then $L$ has a finite minimum representative set.
  Furthermore, letting $\Gamma = \prop{L}$, there exist finite sets
  $\V \subseteq \reals^\Y_+$ and
  $\Theta = \{\theta_v \subseteq \simplex \mid v\in\V\}$,
  both uniquely determined by $\risk{L}$ alone,
  such that
  \begin{enumerate}
  % \item There exists $\V = \V(\risk{L}) \subseteq \reals^\Y_+$, uniquely determined by $\risk{L}$ alone, such that $\V = L(\R^*)$.\label{item:X-V}
  \item A set $\R'\subseteq\R$ is representative if and only if $\V \subseteq L(\R')$.\label{item:X-rep-V}
  \item A set $\R'\subseteq\R$ is minimum representative if and only if $L(\R') = \V$.\label{item:X-min-V}
  \item A set $\R'\subseteq\R$ is representative if and only if $\Theta \subseteq \{\Gamma_r \mid r \in \R'\}$.\label{item:X-rep-Theta}
  \item A set $\R'\subseteq\R$ is minimum representative if and only if $\{\Gamma_r \mid r \in \R'\} = \Theta$.\label{item:X-min-Theta}
  \item Every representative set for $L$ contains a minimum representative set for $L$.\label{item:X-rep-contain-min}
  \item The set of full-dimensional level sets of $\Gamma$ is exactly $\Theta$.\label{item:X-full-dim}
  \item For any $r \in \R$, there exists $\theta \in \Theta$ such that $\Gamma_r \subseteq \theta$.\label{item:X-redundant}
  \item $L$ tightly embeds $\ell:\R'\to\reals^\Y_+$ if and only if $\ell$ is injective and $\ell(\R') = \V$.
  \end{enumerate}
\end{lemma}

@@

Previous work from~\citet[Proposition 4]{duchi2018multiclass} realized the significance of matching Bayes risks for calibration with respect to the 0-1 loss.
Proposition~\ref{prop:embed-bayes-risks} broadens this general insight to any discrete loss.
Moreover, their result relies the Bayes risk of the surrogate being strictly concave, whereas polyhedral Bayes risks are never strictly concave.

@@

Motivated by this observation, let us define the ``trim'' of a loss to be this unique set of loss vectors induced by any minimum representative set, as guaranteed by Lemma~\ref{lem:X}(\ref{item:X-min-V}).
\begin{definition}[Trim]\label{def:trim-loss}
  Given a loss $L:\R \to \reals_+^\Y$ with a finite representative set, we define $\trimcover(L) = \{L(r) \mid r \in \R'\}$ given any minimum representative set $\R'$ for $L$.
\end{definition}

Using this notion of trimming a loss, we can again recast our embedding condition: a loss embeds another if and only if they have the same $\trimcover$.

@@

\begin{corollary}\label{cor:trim-loss-red}
  Let $L : \R \to \reals^\Y_+$ have a finite representative set.
  Then $\trimcover(L) = \{L(u) \mid u \in \R \textrm{ s.t. } \neg\exists u'\in\R,u'\neq u,\, \Gamma_u \subsetneq \Gamma_{u'}\}$. 
\end{corollary}
\begin{proof}
  \raf{Didn't get to this proof}

@@

We conclude this section with a similar structural result about the properties elicited or embedded by a surrogate.
As we define embeddings from one loss to another, one can define embeddings from one property to another, where essentially condition (i) is relaxed; see \S~\ref{app:embed-props} for a precise definition.
When working with properties, it is perhaps more natural to define ``trim'' to remove any level sets which are redundant in the sense of set inclusion, as in Corollary~\ref{cor:trim-loss-red}.
Precisely, given a property $\Gamma:\simplex\toto\R$, define\raft{I think the ``$r'\neq r$'' is not needed, right?}
\[ \trimred(\Gamma) := \{\Gamma_r \mid r \in \R \textrm{ s.t. } \neg\exists r'\in\R,r'\neq r,\, \Gamma_r \subsetneq \Gamma_{r'}\}~.\]

With these definitions, we now state a structural result for embeddable properties, and the properties that embed them.
First, if $\Gamma$ embeds $\gamma$ and $\gamma$ is non-redundant, the level sets of $\Gamma$ must all be redundant relative to $\gamma$.
In other words, $\Gamma$ is exactly the property $\gamma$, just with other reports filling in the gaps between the embedded reports of $\gamma$.
(When working with convex losses, these extra reports are typically the convex hull of the embedded reports.)
In this sense, we can regard embedding as only a slight departure from direct elicitation: if a loss $L$ elicits $\Gamma$ which embeds $\gamma$, we can think of $L$ as essentially eliciting $\gamma$ itself.
Finally, we have an important converse: if $\Gamma$ has finitely many full-dimensional level sets, or if $\trimred(\Gamma)$ is finite, then $\Gamma$ must embed some finite elicitable property with those same level sets.

\begin{restatable}{proposition}{embedtrim}\label{prop:embed-trim}
  Let $\Gamma:\simplex\toto\reals^d$ be an elicitable property.
  The following are equivalent:
  \begin{enumerate}\setlength{\itemsep}{0pt}
  \item $\Gamma$ embeds a elicitable finite property $\gamma:\simplex \toto \R$.
  \item $\trimred(\Gamma)$ is a finite set.%, and $\cup\,\trimred(\Gamma) = \simplex$.
  % \\ $\hat\Gamma := \trimred(\Gamma)$ is a finite property, and $\cup\,\hat\Gamma) = \simplex$.
  \item There is a finite minimum representative set $\U$ for $\Gamma$.
  \item There is a finite set of full-dimensional level sets $\Theta$ of $\Gamma$, and $\cup\,\Theta = \simplex$.
  \end{enumerate}
  % \raft{I see why you wanted this addition here, but (a) it's clear from the def of embedding, and (b) we don't have varphi defined.  So let's leave it out.}
  Moreover, when any of the above hold, $\trimred(\gamma) = \trimred(\Gamma) = \{\Gamma_u \mid u\in\U\} = \Theta$.
\end{restatable}

We defer the proof, and precise definitions, to \S~\ref{app:embed-props}.
